---
title: "tcr_distributions"
author: "Doug McNeall"
date: "2/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Possible distributions for TCR

There are a number of estimates for the pdf of TCR. Here are some possible fits to the stated distributions.



## Quantiles from the literature

There are a number of sources of assessed probability distributions (sometimes just quantiles) in the literature. A complicating factor is that they often give a small number of different quantiles, or probability ranges, leaving a large number of potential distributions that would fit. For example, they might state a 66% probability range, without specifying the quantiles this range apllies to.


### AR5 (assessed range)

"likley" range (66% probability)

1 - 2.5 deg C

Also "positive" and "extremely unlikely" (5% probability) above 3"

I interpret this as

| quantile  |  0% |  16% | 84% | 97.5% |
|-----------|-----|------|-----|-------|
|     tcr   |  0  | 1    | 2.5 |  3    |


But it's possible to interpret that upper quantile as 95%, as there is ambiguity around whether the 5% probability should be split over both tails of the distribution. . I'll update this possible inconsistancy as I find out. 

### CMIP5 models
AR5 Box 12.2 states:

"5 to 95% range of CMIP5 (1.2°C to 2.4°C; see Table 9.5), is positive and extremely unlikely greater than 3°C." As this explicitly mentions the 95% *and* the "extremely unlikely" statement, that lends evidence that "extremely unlikely" is spread over both tails, and refers to the 97.5th percentile.

| quantile  |  0% |  5% | 95% | 97.5% |
|-----------|-----|-----|-----|-------|
|     tcr   |  0  | 1.2 | 2.5 |  3    |


### Schurer eta al (2018)

Multimodel mean with increased variance for model uncertainty

| quantile  |  5% | 50% | 95% |
|-----------|-----|-----|-----|
|     tcr   | 1.2 | 1.7 | 2.4 |


### Sherwood et al (2020)

Likely (66% probability) range

| quantile  | 16% | 50% | 84% |
|-----------|-----|-----|-----|
|     tcr   | 1.5 | 1.8 | 2.2 |


### Jiménez-de-la Cuesta and Mauritsen (2019)

| quantile  |  5%  | 50% | 95%  |
|-----------|------|-----|------|
|     tcr   | 1.17 | 1.7 | 2.16 |


### Tokarska et al (2020)

| quantile  | 16% | 50% | 84% |
|-----------|-----|-----|-----|
|     tcr   | 1.2 | 1.6 | 2.0 |


### Njisse et al. (2020)

Njisse et al. (2020) CMIP5 models

| quantile  |  5%  |  16%  |  50% | 84%  |  95%  |
|-----------|------|-------|------|------|-------|
|     tcr   | 1.1  | 1.4   | 1.7  |  2.1 |  2.4  |

Njisse et al. (2020) CMIP6 models

| quantile  |  5%  |  16%  |  50% | 84%  |  95%  |
|-----------|------|-------|------|------|-------|
|     tcr   |   1  | 1.29  | 1.68 | 2.05 |  2.3  |


## Distributions used in Betts & McNeall (2018)

```{r}
dx = 100
tcr_range = seq(from = 0, to = 5, by = 1/dx)

```

```{r}
# Very approximately, the TCR distribution is N(1.75, 0.5)
# MC sample, and plug that into the transfer function
# Still to do: pick a closer mean/sd (perhaps non normal?)
tcrmean = 1.75
tcrsd_used = 0.45 # incorrectly used in the paper.
tcrsd = 0.75 # The final result is very sensitive to lower tail of tcr distribution

qnorm(p = c(0.16, 0.84), mean = tcrmean, sd = tcrsd_used)
qnorm(p = c(0.16, 0.84), mean = tcrmean, sd = tcrsd)

# A plot of what the assumed tcr distribution looks like
tcrdens_used = dnorm(tcr_range, mean = tcrmean, sd = tcrsd_used)
#tcrsamp_used  = rnorm(10000, mean = tcrmean, sd = tcrsd_used)

tcrdens = dnorm(tcr_range, mean = tcrmean, sd = tcrsd)
par(las = 1)
plot(tcr_range, tcrdens_used, type = 'l', lwd = 2, main = "pr(TCR)", xlab = "TCR", ylab = "Probability density" )
lines(tcr_range, tcrdens, col = 'red', lwd = 2 )
legend('topright', c('used in paper', 'corrected Normal distribution'), lty = 'solid', col = c('black', 'red'))
```



## Other distributions to might fit TCR better

The distribution from Richardson is asymmetric, and there is evidence that the other distributions representing scientists beliefs are asymmetric too. Normal distributions can easily place significant probability below zero, which is explicitly ruled out in the AR5 assessment.

Other distributions, such as Gamma or lognormal, might fit better.

For example, here is a Gamma distribution with some sensible-looking parameters.

```{r}

gdens <- dgamma(x = tcr_range, shape = 5, rate = 3)

plot(tcr_range, gdens, type = 'l', lwd = 2, main = "pr(TCR)", xlab = "TCR", ylab = "Probability density" )

```

## Fit distributions to the Richardson samples from their TCR distribution


```{r}
# Load Richardson TCR samples from distribution
atcrpdf = as.numeric(readLines("../richardson/atcrpdf.txt"))
atcrpdf.trunc = atcrpdf[atcrpdf<12 & atcrpdf >0]
```

```{r}
library(fitdistrplus)
```

Fit the Gamma distribution, starting from the above parameters.

```{r}
start.gamma <- list(shape = 5, rate = 3)
fit.gamma <- fitdist(atcrpdf.trunc, distr = "gamma", method = "mle", start = start.gamma)
summary(fit.gamma)

```

```{r}

plot(fit.gamma)

```

### Fit a lognormal distribution to Richardson

This seems to fit a little better than the Gamma distribution.


```{r}
# lognormal?
# A lognormal distribution seems to fit the 

fit.lnorm <- fitdist(atcrpdf.trunc, distr = "lnorm", method = "mle")
summary(fit.lnorm)

plot(fit.lnorm)

exp(fit.lnorm$estimate)

```

## Fit a lognormal distribution to all of the distributions from the literature

```{r}

library(rriskDistributions)

# Quantiles and probabilities from the various estimates

# AR5 assessed likely
p_ar5 <- c(0.16, 0.84, 0.975)
q_ar5 <- c(1, 2.5, 3)

# CMIP5 models
p_cmip5 <- c(0.05, 0.95, 0.975)
q_cmip5 <- c(1.2, 2.5, 3)

p_schurer <- c(0.05,0.50, 0.95)
q_schurer <- c(1.2, 1.7, 2.4)

p_sherwood <- c(0.16,0.50, 0.84)
q_sherwood <- c(1.5, 1.8, 2.2)

p_richardson <- c(0.05, 0.50, 0.95)
q_richardson <- quantile(atcrpdf.trunc, p_richardson) # check these

p_njisse_cmip5 <- c(0.05, 0.16, 0.5, 0.84, 0.95)
q_njisse_cmip5 <- c(1.1, 1.4, 1.7, 2.1, 2.4 )

p_njisse_cmip6 <- c(0.05, 0.16, 0.5, 0.84, 0.95)
q_njisse_cmip6 <- c(1, 1.29, 1.68, 2.05, 2.3)
# Calculate the parameters of a lognormal distribution that would fit those parameters


p_jimenez <- c(0.05, 0.5, 0.95)
q_jimenez <- c(1.17, 1.7, 2.16)
  
p_tokarska <- c(0.16, 0.5, 0.84)
q_tokarska <- c(1.2, 1.6, 2.0)
  

lnpar_ar5 <- get.lnorm.par(p = p_ar5, q_ar5) 

lnpar_cmip5 <- get.lnorm.par(p=p_cmip5 , q=q_cmip5)

# Schurer paper
lnpar_schurer <- get.lnorm.par(p=p_schurer, q=q_schurer)

# Sherwood paper
lnpar_sherwood <- get.lnorm.par(p=p_sherwood, q=q_sherwood)

#richardson paper
# This is different from the fit.lnorm
lnpar_richardson <- get.lnorm.par(p=p_richardson, q=q_richardson)

lnpar_njisse_cmip5 <- get.lnorm.par(p=p_njisse_cmip5, q = q_njisse_cmip5)

lnpar_njisse_cmip6 <- get.lnorm.par(p=p_njisse_cmip6, q = q_njisse_cmip6)

lnpar_jimenez <- get.lnorm.par(p=p_jimenez, q = q_jimenez)

lnpar_tokarska<- get.lnorm.par(p=p_tokarska, q = q_tokarska)


```


## Plot all of the lognormal fits

```{r, fig.width = 10, fig.height =7}

ar5_ln_dens <- dlnorm(x = tcr_range, meanlog = lnpar_ar5['meanlog'], sdlog = lnpar_ar5['sdlog'] )
cmip5_ln_dens <- dlnorm(x = tcr_range, meanlog = lnpar_cmip5['meanlog'], sdlog = lnpar_cmip5['sdlog'] )
schurer_ln_dens <- dlnorm(x = tcr_range, meanlog = lnpar_schurer['meanlog'], sdlog = lnpar_schurer['sdlog'] )
sherwood_ln_dens <- dlnorm(x = tcr_range, meanlog = lnpar_sherwood['meanlog'], sdlog = lnpar_sherwood['sdlog'] )
richardson_ln_dens <- dlnorm(x = tcr_range, meanlog = lnpar_richardson['meanlog'], sdlog = lnpar_richardson['sdlog'])
njisse_cmip5_ln_dens <- dlnorm(x = tcr_range, meanlog = lnpar_njisse_cmip5['meanlog'], sdlog = lnpar_njisse_cmip5['sdlog'])  
njisse_cmip6_ln_dens <- dlnorm(x = tcr_range, meanlog = lnpar_njisse_cmip6['meanlog'], sdlog = lnpar_njisse_cmip6['sdlog'])
jimenez_ln_dens <- dlnorm(x = tcr_range, meanlog = lnpar_jimenez['meanlog'], sdlog = lnpar_jimenez['sdlog'])
tokarska_ln_dens <- dlnorm(x = tcr_range, meanlog = lnpar_tokarska['meanlog'], sdlog = lnpar_tokarska['sdlog'])

h <- hist(atcrpdf.trunc, breaks = 30, plot = FALSE)

yfit_richardson <- richardson_ln_dens*diff(h$mids[1:2])*length(atcrpdf.trunc)
yfit_schurer <- schurer_ln_dens*diff(h$mids[1:2])*length(atcrpdf.trunc)
yfit_sherwood <- sherwood_ln_dens*diff(h$mids[1:2])*length(atcrpdf.trunc)
yfit_cmip5 <- cmip5_ln_dens*diff(h$mids[1:2])*length(atcrpdf.trunc)
yfit_ar5 <- ar5_ln_dens*diff(h$mids[1:2])*length(atcrpdf.trunc)
yfit_njisse_cmip5 <- njisse_cmip5_ln_dens*diff(h$mids[1:2])*length(atcrpdf.trunc)
yfit_njisse_cmip6 <- njisse_cmip6_ln_dens*diff(h$mids[1:2])*length(atcrpdf.trunc)
yfit_jimenez <- jimenez_ln_dens*diff(h$mids[1:2])*length(atcrpdf.trunc)
yfit_tokarska <- tokarska_ln_dens*diff(h$mids[1:2])*length(atcrpdf.trunc)

rn <- range(yfit_richardson, yfit_schurer,yfit_sherwood,yfit_cmip5, yfit_ar5,yfit_njisse_cmip5, yfit_njisse_cmip6,yfit_jimenez,yfit_tokarska)

plot(h, xlim = c(0,4), ylim = rn, xlab = 'TCR', ylab = 'density', main = 'lognormal fits to quantiles')
lines(tcr_range, yfit_richardson, lwd = 3, col = 'orange')
lines(tcr_range, yfit_schurer, lwd = 3, col = 'red')
lines(tcr_range, yfit_sherwood, lwd = 3, col = 'blue')
lines(tcr_range, yfit_cmip5, lwd = 3, col = 'black')
lines(tcr_range, yfit_ar5, lwd = 3, col = 'pink')
lines(tcr_range, yfit_njisse_cmip5, lwd = 3, col = 'purple', lty = 'dashed')
lines(tcr_range, yfit_njisse_cmip6, lwd = 3, col = 'purple')
lines(tcr_range, yfit_jimenez, lwd = 3, col = 'darkgrey')
lines(tcr_range, yfit_tokarska, lwd = 3, col = 'darkblue')

legend('topright', c('Richardson 16', 'Schurer 18', 'Sherwood 20', 'CMIP5 13', 'AR5 13', 'Njisse CMIP5', 'Njisse CMIP6', 'Jimenez 2019', 'Tokarska 2020'),
       col = c('orange', 'red', 'blue', 'black','pink', 'purple', 'purple', 'darkgrey', 'darkblue'), lwd = 3, lty = c('solid','solid','solid','solid','solid','dashed','solid', 'solid', 'solid'))


```


## How close are the lognormal fits?
This section plots the fitted quantiles (on the y axis) against the given quantiles (from the literature, on the x axis). Lower quantiles (with values below 3) are in general well fitted. The fitted upper quantile of the distribution is much higher than suggested by the given distribution, suggesting that the fit assigns too much probability to the upper tail of the distribution than the assesed distribution would suggest is right.

```{r, fig.width = 6, fig.height = 6}

# Plot theoretical vs actual quantile for each fit
par(las = 1)
plot(q_ar5, qlnorm(p=p_ar5, meanlog = lnpar_ar5['meanlog'],  sdlog =  lnpar_ar5['sdlog']),
     col = 'pink', pch = 19,
     xlim = c(0.5,4.5), ylim = c(0.5,4.5),
     xlab = 'given quantile', ylab = 'fitted quantile')
points(q_cmip5, qlnorm(p=p_cmip5, meanlog = lnpar_cmip5['meanlog'],  sdlog =  lnpar_cmip5['sdlog']),
       col = 'black', pch = 19)
points(q_sherwood, qlnorm(p=p_sherwood, meanlog = lnpar_sherwood['meanlog'],  sdlog =  lnpar_sherwood['sdlog']),
       col = 'blue', pch = 19)
points(q_schurer, qlnorm(p=p_schurer, meanlog = lnpar_schurer['meanlog'],  sdlog =  lnpar_schurer['sdlog']),
       col = 'red', pch = 19)
points(q_richardson, qlnorm(p=p_richardson, meanlog = lnpar_richardson['meanlog'],  sdlog =  lnpar_richardson['sdlog']),
       col ='orange', pch = 19)
legend('topleft', c('Richardson 16', 'Schurer 18', 'Sherwood 20', 'CMIP5 13', 'AR5 13'),
       col = c('orange', 'red', 'blue', 'black','pink'), pch = 19)

abline(0,1)




```





```{r}

h <- hist(atcrpdf.trunc, prob = TRUE, ylim = c(0,1), xlim = c(0,4), main = 'Fit to richardson distribution')
d <- density(atcrpdf.trunc)

lines(d, lwd = 2)
lines(tcr_range,richardson_ln_dens, col = 'orange', lwd = 2)
legend('topright', legend = c('density estimate', 'lognormal fit'), col = c('black', 'orange'), lwd = 2)


```



